Measurement and Probability in Fuzzy Quantum Theories

### 2016-11-24 proposal/prop.tex

In a series of breakthrough papers that led to his Turing
Award~\cite{Milner:1993:EIT:151233.151240}, Milner argued that this
``closed system'' view of computation is too limiting as it cannot
adequately describe \emph{interactive computations}. Milner explained
that ``concurrency requires a new conceptual framework'' and ``not
merely an extension of the repertoire of entities and constructions
which explain sequential computing.'' He gives the following simple
but compelling example. Consider the following two programs:
\begin{verbatim}
Program P1: x := 1 ; x := x + 1 
Program P2: x := 2
\end{verbatim}
As closed systems, these two programs define the same relation, namely
the one that maps an input memory to an updated memory in which the
location named \verb|x| is assigned the value \verb|2|. But place the
programs in an environment which may involve interactions and the
equivalence is lost. Concretely, place each of the two programs
concurrently with the program:
\begin{verbatim}
Program Q: x := 3
\end{verbatim}
and we can observe the following. In the combined system of \verb|P1|
and \verb|Q| running concurrently, the location \verb|x| may end up
containing \verb|2|, \verb|3|, or \verb|4|. But in the combined system
of \verb|P2| and \verb|Q| running concurrently, the location~\verb|x|
can only end up containing \verb|2| or \verb|3|. The lesson, phrased
in physics terms, is that it not sensible to think of the properties
of a system (like \verb|P1| or \verb|P2|) independently of the
observer (\verb|Q| in this case).

@@

\item If we insist that probabilities cannot be infinitely precise and
  are bound to be approximations represented by intervals of
  confidence, then by Gleason's theorem, quantum states themselves
  cannot be infinitely precise. This raises an important philosophical
  question with immense practical relevance in the context of quantum
  computing. The question, which we could largely bypass in the
  classical case, is whether there is, in the background of all the
  measurements and probabilities, a ``real'' infinitely-precise
  quantum state that is being approximated and that, in principle, can
  be more and more accurately described given more and more
  resources. The answer will have to be \emph{no} if we are to respect
  major theorems of quantum mechanics and quantum information. Thus
  our first radical departure from the classical case will be that the
  quantum state itself must be fuzzy in some sense to be formalized.

>>>>

In a series of breakthrough papers that led to his Turing
Award~\cite{Milner:1993:EIT:151233.151240}, Milner argued that this
``closed system'' view of computation is too limiting as it cannot
adequately describe \emph{interactive computations}. Milner explained
that ``concurrency requires a new conceptual framework'' and ``not
merely an extension of the repertoire of entities and constructions
which explain sequential computing.'' He gives the following simple
but compelling example. Consider the following two programs:
\begin{verbatim}
Program P1: x := 1 ; x := x + 1 
Program P2: x := 2
\end{verbatim}
As closed systems, these two programs define the same relation, namely
the one that maps an input memory to an updated memory in which the
location named \verb|x| is assigned the value \verb|2|. But place the
programs in an environment which may involve interactions and the
equivalence is lost. Concretely, place each of the two programs
concurrently with the program:
\begin{verbatim}
Program Q: x := 3
\end{verbatim}
and we can observe the following. In the combined system of \verb|P1|
and \verb|Q| running concurrently, the location \verb|x| may end up
containing \verb|2|, \verb|3|, or \verb|4|. But in the combined system
of \verb|P2| and \verb|Q| running concurrently, the location~\verb|x|
can only end up containing \verb|2| or \verb|3|. The lesson, phrased
in physics terms, is that it is not sensible to think of the properties
of a system (like \verb|P1| or \verb|P2|) independently of the
observer (\verb|Q| in this case). Indeed the observation \verb|x=4|
cannot said to have an independent existence independently from the
interaction between \verb|P1| and \verb|Q|: it really emerges or is
created as part of the observation process and is not a pre-existing
``element of reality''.

@@

\item If we insist that probabilities cannot be infinitely precise and
  are bound to be approximations represented by intervals of
  confidence, then by Gleason's theorem, quantum states themselves
  cannot be infinitely precise. This raises an important philosophical
  question with immense practical relevance in the context of quantum
  computing. The question, which we could largely bypass in the
  classical case, is whether there is, in the background of all the
  measurements and probabilities, a ``real'' infinitely-precise
  quantum state that is being approximated and that, in principle, can
  be more and more accurately described given more and more
  resources. The answer will have to be \emph{no} if we are to respect
  major theorems of quantum mechanics and quantum information. Thus
  our first radical departure from the classical case will be that the
  quantum state itself must be fuzzy in some sense to be
  formalized. In other words, does a Gleason, Bell, or Kochen-Specker
  theorem still hold in a world with finite resources?
